HealpixNet: Efficient spherical Convolutional Neural Network with Healpix sampling for cosmological applications

### 2018-07-17 scnn.tex
figures: omit path and extension

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{color}
\usepackage{url}
\usepackage{hyperref}
\usepackage{graphicx,array}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{bm}
\usepackage{aasmacros}

@@

\begin{figure*}[!ht]
	\centering
    % \includegraphics[width=\linewidth]{figures/architecture}
	\includegraphics[width=\linewidth]{figures/figure_architecture.pdf}
	\caption{Overall network architecture, showing here two convolutional layers acting as feature extractors followed by a fully connected layer with softmax acting as the classifier. A convolutional layer is based on five operations: convolution, non-linearity, batch normalization, down-sampling, and pooling. While most operations are agnostic to the data domain, the convolution and the down-sampling have to be adapted. In this paper, we propose first to model the sphere with a graph and to perform the convolution on the graph. Graphs are versatile data structures which can model any sampling, even irregular or partial. Second, we propose to exploit a hierarchical pixelization of the sphere for the pooling operation. It allows the network to analyze the data at multiple scales while preserving the spatial localization of features.
    \nati{Great plot! Could you make the sphere round? If you use the current code in paper plot notebook, they will be.}
    }
	\label{fig:architecture}
\end{figure*}

@@

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exampleEarthTopo}
		\caption{}
		\label{fig:example_earth}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exampleCMB}
		\caption{}
		\label{fig:example_cmb}
	\end{subfigure}
	\caption[]{Example of (a) an Earth topography map composed of \textasciitilde 3 million pixels (\textasciitilde 7 arcmin resolution) and (b) a model of the Cosmic Microwave Background (CMB) radiation temperature anisotropy, composed of \textasciitilde 13 million pixels (\textasciitilde 3.4 arcmin resolution).\footnotemark}
	\label{fig:example_maps}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{figures/half_graph_4.pdf}
	\caption[]{Graph built from an HEALPix pixelization of half the sphere ($N_{side} = 4$). By construction, each node has 8 neighbors, except the highlighted ones which have only 7.\footnotemark[6]}
	\label{fig:healpix_graph_4}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{figures/gaussian_filters_sphere.pdf}
	\caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
	Top: Diffusion of a unit of heat for different times $t$ using the graph.
	Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
	Relative difference between graph convolution and spherical smoothing: $10.4$\%, $5.7$\%, $4.8$\%, $3.8$\% }
	\label{fig:gaussian_filters_comparizon}
\end{figure*}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{figures/filtering_speed}
	\caption{Comparison of filtering speed for Gaussian smoothing of maps of various sizes. The fast spherical harmonic transform (SHT) is implemented by the healpy Python package (via the \texttt{healpy.sphtfunc.smoothing} function with $\ell_{max} = 3 N_{side} - 1$, the default.). The graph filtering is defined by \eqnref{graph_convolution_cheby} and implemented with the numpy and scipy Python packages. Both are executed on a single core. The theoretical cost of filtering on the graph is $\bO(K N_{pix})$ and $\bO(\ell_{max}^3)$ for the spherical harmonics, where $\ell_{max}$ is the largest degree (i.e., the spectrum is truncated at $\ell_{max}$).}
	\label{fig:filtering_speed}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{figures/pooling.pdf}
	\caption{Two levels of coarsening and pooling: groups of 4 cells are merged into one, then the data on them is summarized in one. The coarsest cell covers $1/12$ of the sphere.}
	\label{fig:pooling}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/psd_sigma3.pdf}
\caption{Power spectrum densities of the noiseless maps. The maps have been smoothed with a Gaussian kernel of radius $3$ arcmins to remove high frequencies. The cosmological models could otherwise have been distinguished by looking at the PSDs.}
\label{fig:psd_sigma3}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.48\linewidth]{figures/smooth_map_class_1.pdf}
\hfill
\includegraphics[width=0.48\linewidth]{figures/smooth_map_class_2.pdf}
\caption{Example maps from two classes to be discriminated. Left: model 1 with $\Omega_m=0.31$ and $\sigma_8=0.82$. Right: model 2 with $\Omega_m=0.26$ and $\sigma_8=0.91$.
The initial conditions for both simulations were the same, so differences only arise due to different cosmological parameters. \nati{Tomek: Is this correct?}}
\label{fig:map_sample}
\end{figure*}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/part_sphere.pdf}
\caption{Three parts of the sphere with different sizes. Blue: $o=1$. Green: order $o=2$. Yellow: order $o=4$.}
\label{fig:part_sphere}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.32\linewidth]{figures/result_order1.pdf}
\hfill
\includegraphics[width=0.32\linewidth]{figures/result_order2.pdf}
\hfill
\includegraphics[width=0.32\linewidth]{figures/result_order4.pdf}
\caption{Classification errors for the 3 different problems.}
\label{fig:results}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.47\linewidth]{figures/section_filter_last.pdf}
\hfill
\includegraphics[width=0.45\linewidth]{figures/gnonomic_filter_last.pdf}
\caption{24 learned filters selected from the last spherical convolutional layer. Left: section. Right: gnonomic view.}
\label{fig:learned_filter}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/gaussian_filters_spectral.pdf}
\includegraphics[width=\linewidth]{figures/gaussian_filters_gnomonic.pdf}
\includegraphics[width=\linewidth]{figures/gaussian_filters_section.pdf}
\caption{Three different visualizations of the convolution kernel $K_t(x)=e^{-\tau t x}$.
Top: graph spectral domain.
Middle: gnomonic projection.
Bottom: section plot.}
\label{fig:gaussian_filters_visualization}
\end{figure*}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.6\linewidth]{figures/index_plotting_order20_nside16.pdf}
\caption{Indexes selected for the section ploting of \figref{gaussian_filters_visualization} middle. The delta was placed on the yellow node.}
\label{fig:index_section}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{figures/hist_error_order2_noise1_5.pdf}
\caption{Evolution curve of the error with respect to the number of training samples for histogram features and a linear classifier (setting: order 2, relative noise level: 1.5). The validation error might be slightly below the training error because we use it for cross-validation.}
\label{fig:hist_error_evolution}
\end{figure}

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{color}
\usepackage{url}
\usepackage{hyperref}
\usepackage{graphicx,array} \graphicspath{{figures/}}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{siunitx}
\usepackage{bm}
\usepackage{aasmacros}

@@

\begin{figure*}
	\centering
    % \includegraphics[width=\linewidth]{architecture}
	\includegraphics[width=\linewidth]{figure_architecture}
	\caption{Overall network architecture, showing here two convolutional layers acting as feature extractors followed by a fully connected layer with softmax acting as the classifier. A convolutional layer is based on five operations: convolution, non-linearity, batch normalization, down-sampling, and pooling. While most operations are agnostic to the data domain, the convolution and the down-sampling have to be adapted. In this paper, we propose first to model the sphere with a graph and to perform the convolution on the graph. Graphs are versatile data structures which can model any sampling, even irregular or partial. Second, we propose to exploit a hierarchical pixelization of the sphere for the pooling operation. It allows the network to analyze the data at multiple scales while preserving the spatial localization of features.
    \nati{Great plot! Could you make the sphere round? If you use the current code in paper plot notebook, they will be.}
    }
	\label{fig:architecture}
\end{figure*}

@@

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exampleEarthTopo}
		\caption{}
		\label{fig:example_earth}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.45\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exampleCMB}
		\caption{}
		\label{fig:example_cmb}
	\end{subfigure}
	\caption[]{Example of (a) an Earth topography map composed of \textasciitilde 3 million pixels (\textasciitilde 7 arcmin resolution) and (b) a model of the Cosmic Microwave Background (CMB) radiation temperature anisotropy, composed of \textasciitilde 13 million pixels (\textasciitilde 3.4 arcmin resolution).\footnotemark}
	\label{fig:example_maps}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{half_graph_4}
	\caption[]{Graph built from an HEALPix pixelization of half the sphere ($N_{side} = 4$). By construction, each node has 8 neighbors, except the highlighted ones which have only 7.\footnotemark[6]}
	\label{fig:healpix_graph_4}
\end{figure}

@@

\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{gaussian_filters_sphere}
	\caption{Comparison of convolution with the graph and the spherical harmonics ($N_{side} = 16$).
	Top: Diffusion of a unit of heat for different times $t$ using the graph.
	Bottom: spherical symmetric Gaussian smoothing for different $\sigma$ (arcmin).
	Relative difference between graph convolution and spherical smoothing: $10.4$\%, $5.7$\%, $4.8$\%, $3.8$\% }
	\label{fig:gaussian_filters_comparizon}
\end{figure*}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{filtering_speed}
	\caption{Comparison of filtering speed for Gaussian smoothing of maps of various sizes. The fast spherical harmonic transform (SHT) is implemented by the healpy Python package (via the \texttt{healpy.sphtfunc.smoothing} function with $\ell_{max} = 3 N_{side} - 1$, the default.). The graph filtering is defined by \eqnref{graph_convolution_cheby} and implemented with the numpy and scipy Python packages. Both are executed on a single core. The theoretical cost of filtering on the graph is $\bO(K N_{pix})$ and $\bO(\ell_{max}^3)$ for the spherical harmonics, where $\ell_{max}$ is the largest degree (i.e., the spectrum is truncated at $\ell_{max}$).}
	\label{fig:filtering_speed}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{pooling}
	\caption{Two levels of coarsening and pooling: groups of 4 cells are merged into one, then the data on them is summarized in one. The coarsest cell covers $1/12$ of the sphere.}
	\label{fig:pooling}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{psd_sigma3}
\caption{Power spectrum densities of the noiseless maps. The maps have been smoothed with a Gaussian kernel of radius $3$ arcmins to remove high frequencies. The cosmological models could otherwise have been distinguished by looking at the PSDs.}
\label{fig:psd_sigma3}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.48\linewidth]{smooth_map_class_1}
\hfill
\includegraphics[width=0.48\linewidth]{smooth_map_class_2}
\caption{Example maps from two classes to be discriminated. Left: model 1 with $\Omega_m=0.31$ and $\sigma_8=0.82$. Right: model 2 with $\Omega_m=0.26$ and $\sigma_8=0.91$.
The initial conditions for both simulations were the same, so differences only arise due to different cosmological parameters. \nati{Tomek: Is this correct?}}
\label{fig:map_sample}
\end{figure*}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{part_sphere}
\caption{Three parts of the sphere with different sizes. Blue: $o=1$. Green: order $o=2$. Yellow: order $o=4$.}
\label{fig:part_sphere}
\end{figure}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.32\linewidth]{result_order1}
\hfill
\includegraphics[width=0.32\linewidth]{result_order2}
\hfill
\includegraphics[width=0.32\linewidth]{result_order4}
\caption{Classification errors for the 3 different problems.}
\label{fig:results}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=0.47\linewidth]{section_filter_last}
\hfill
\includegraphics[width=0.45\linewidth]{gnonomic_filter_last}
\caption{24 learned filters selected from the last spherical convolutional layer. Left: section. Right: gnonomic view.}
\label{fig:learned_filter}
\end{figure*}

@@

\begin{figure*}[!ht]
\centering
\includegraphics[width=\linewidth]{gaussian_filters_spectral}
\includegraphics[width=\linewidth]{gaussian_filters_gnomonic}
\includegraphics[width=\linewidth]{gaussian_filters_section}
\caption{Three different visualizations of the convolution kernel $K_t(x)=e^{-\tau t x}$.
Top: graph spectral domain.
Middle: gnomonic projection.
Bottom: section plot.}
\label{fig:gaussian_filters_visualization}
\end{figure*}

\begin{figure}[!ht]
\centering
\includegraphics[width=0.6\linewidth]{index_plotting_order20_nside16}
\caption{Indexes selected for the section ploting of \figref{gaussian_filters_visualization} middle. The delta was placed on the yellow node.}
\label{fig:index_section}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\linewidth]{hist_error_order2_noise1_5}
\caption{Evolution curve of the error with respect to the number of training samples for histogram features and a linear classifier (setting: order 2, relative noise level: 1.5). The validation error might be slightly below the training error because we use it for cross-validation.}
\label{fig:hist_error_evolution}
\end{figure}
